Estimating the dynamics of SPARQL query results

### 2019-04-11 kgswc19.tex
update

\begin{document}
%
\title{Estimating the dynamics of SPARQL query results
%\thanks{This work was supported by CONICYT PFCHA, Doctorado Becas Chile/2017 -- 72180000, the Millennium Institute for Foundational Research on Data (IMFD) and FONDECYT Grant No. 1181896.}
}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
%\author{Alberto Moya\orcidID{0000-0002-7003-5087} \and 
%Aidan Hogan\orcidID{0000-0001-9482-1982}}
%
%\authorrunning{A. Moya and A. Hogan}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
%\institute{IMFD; DCC, University of Chile\\
%\email{\{amoya,ahogan\}@dcc.uhile.cl}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
We address the problem of estimating when the results of an input SPARQL query results over dynamic RDF datasets will change. We propose a framework that extracts features from the query and/or from past versions of the target dataset. The resulting features are then used to predict (1) whether or not the results for a query will change in the near future using classifiers, and (2) when they might change (if at all) using linear regression. To evaluate our proposal, we create a gold standard based on 23 versions of Wikidata and a curated collection of 221 SPARQL queries. Our results show that the quality of predictions using query features and lightweight statistics of the dynamics of individual predicates is competitive with that obtained using (more costly to derive) knowledge of the complete historical changes in the query results.

@@

In this work, we address the problem of estimating if and/or when SPARQL query results will change in a future version of a dynamic RDF dataset. We propose a general architecture based on Machine Learning that extracts static features from a query, as well as dynamic features from the query and dataset. These features are fed into a classifier to predict whether or not the query results will change in a fixed point in the future, or, more ambitiously, can be fed into a regression model to predict when the query results are likely to change. With respect to the features used, we propose that there is a trade-off between those that are coarse-grained but easy to compute (e.g., static query features) versus those that are fine-grained but more costly to compute (e.g., historical changes in query results), with further features falling in the middle (e.g., high-level statistics about the dynamics of triples with particular predicates). Per this trade-off, which features to use for predicting the dynamics of query results may then depend on the particular application.

In order to better understand this conceptual trade-off -- and more generally, to evaluate the quality of predictions made by our framework -- we create a novel gold standard based on 23 weekly versions of the Wikidata knowledge graph~\cite{VrandecicK14} and 221 user-generated queries; we show this gold standard to have a variety of desirable features, including (most importantly) a balance of queries whose results never change, always change, as well as non-trivial queries whose results intermittently change. Using this gold standard, our experiments show that while the most accurate predictions are based on historical changes to the input query's results, predictions based only on query features and predicate dynamics remain competitive (while being much more efficient to maintain/extract).

\paragraph{Contributions:} (1) We propose a general architecture for predicting when/if the results of a SPARQL query will change in a future version of a dynamic RDF dataset. (2) We propose a number of features to instantiate this architecture based on analysis of the query, analysis of the dynamics of predicates in the data, and analysis of historical changes in the query results. (3) We create a gold standard for these tasks based on 23 consecutive versions of Wikidata and a set of 221 real-world SPARQL queries. (4) We use this gold standard to compare the predictions obtained using different types of features and classifiers.

\paragraph{Paper structure:} In Section~\ref{sec:related}, we review related works. Section~\ref{sec:preliminar} provides a formal statement of the problem. Section~\ref{sec:approach} describes our proposed architecture and features. In Section~\ref{sec:data}, we describe the creation of our gold standard based on Wikidata. Section~\ref{sec:eval} presents the design and results of the experiments. We conclude and review future directions in Section~\ref{sec:conclusion}.

@@

\paragraph{Novelty:} Given a query and a dynamic RDF dataset, we aim to predict (1) whether or not the query's results will have changed in a fixed point in the near future and/or (2) when the results will change. Our work thus complements existing works aiming for weak consistency, but (i) generalizes the problem, proposing a framework that can incorporate statistics on predicate dynamics~\cite{UmbrichKHP12,UmbrichKPPH12,ekawUmbrichKHP12,DehghanzadehPKUHD14} and historical changes in query results~\cite{KnuthHS16}, as proposed in prior works, as well as novel types of features, (ii) introduces new features based on query operators and statistics; (iii) creates a novel gold standard based on Wikidata and presents comparative results that indicate the relative predictive power inherent in different types of features.

>>>>

\begin{document}
%
\title{Estimating the dynamics of SPARQL query results
%\thanks{This work was supported by CONICYT PFCHA, Doctorado Becas Chile/2017 -- 72180000, the Millennium Institute for Foundational Research on Data (IMFD) and FONDECYT Grant No. 1181896.}
}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
%\author{Alberto Moya\orcidID{0000-0002-7003-5087} \and 
%Aidan Hogan\orcidID{0000-0001-9482-1982}}
%
%\authorrunning{A. Moya and A. Hogan}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
%\institute{IMFD; DCC, University of Chile\\
%\email{\{amoya,ahogan\}@dcc.uhile.cl}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
We address the problem of estimating when the results of an input SPARQL query results over dynamic RDF datasets will change. We evaluate a framework that extracts features from the query and/or from past versions of the target dataset. The resulting features are then used to predict (1) whether or not the results for a query will change in the near future using classifiers, and (2) when they might change (if at all) using linear regression. To evaluate, we create a gold standard based on 23 versions of Wikidata and a curated collection of 221 SPARQL queries. Our results show that the quality of predictions using query features and lightweight statistics of the dynamics of individual predicates is less competitive with that obtained using (more costly to derive) knowledge of the complete historical changes in the query results.

@@

In this work, we address the problem of estimating if and/or when SPARQL query results will change in a future version of a dynamic RDF dataset. We evaluate a general architecture based on Machine Learning that extracts static features from a query, as well as dynamic features from the query and dataset. These features are fed into a classifier to predict whether or not the query results will change in a fixed point in the future, or, more ambitiously, can be fed into a regression model to predict when the query results are likely to change. With respect to the features used, we show that there is a trade-off between those that are coarse-grained but easy to compute (e.g., static query features) versus those that are fine-grained but more costly to compute (e.g., historical changes in query results), with further features falling in the middle (e.g., high-level statistics about the dynamics of triples with particular predicates). Per this trade-off, which features to use for predicting the dynamics of query results may then depend on the particular application.

In order to better understand this conceptual trade-off -- and more generally, to evaluate the quality of predictions made by our framework -- we create a novel gold standard based on 23 weekly versions of the Wikidata knowledge graph~\cite{VrandecicK14} and 221 user-generated queries; we show this gold standard to have a variety of desirable features, including (most importantly) a balance of queries whose results never change, always change, as well as non-trivial queries whose results intermittently change. Using this gold standard, our experiments show that while the most accurate predictions are based on historical changes to the input query's results, predictions based only on query features and predicate dynamics remain less competitive (while being much more efficient to maintain/extract).

\paragraph{Contributions:} (1) We evaluate a general architecture for predicting when/if the results of a SPARQL query will change in a future version of a dynamic RDF dataset. (2) We evaluate a number of features to instantiate this architecture based on analysis of the query, analysis of the dynamics of predicates in the data, and analysis of historical changes in the query results. (3) We create a gold standard for these tasks based on 23 consecutive versions of Wikidata and a set of 221 real-world SPARQL queries. (4) We use this gold standard to compare the predictions obtained using different types of features and classifiers.

\paragraph{Paper structure:} In Section~\ref{sec:related}, we review related works. Section~\ref{sec:preliminar} provides a formal statement of the problem. Section~\ref{sec:approach} describes the evaluated architecture and features. In Section~\ref{sec:data}, we describe the creation of our gold standard based on Wikidata. Section~\ref{sec:eval} presents the design and results of the experiments. We conclude and review future directions in Section~\ref{sec:conclusion}.

@@

\paragraph{Novelty:} Given a query and a dynamic RDF dataset, we aim to predict (1) whether or not the query's results will have changed in a fixed point in the near future and/or (2) when the results will change. Our work thus complements existing works aiming for weak consistency, but (i) generalizes the problem, evaluating a framework that can incorporate statistics on predicate dynamics~\cite{UmbrichKHP12,UmbrichKPPH12,ekawUmbrichKHP12,DehghanzadehPKUHD14} and historical changes in query results~\cite{KnuthHS16}, as proposed in prior works, as well as novel types of features, (ii) introduces new features based on query operators and statistics; (iii) creates a novel gold standard based on Wikidata and presents comparative results that indicate the relative predictive power inherent in different types of features.
